\LARGE \bf Human-Robot Interaction: Tackling the AI Challenges

### 2014-05-06 paper.tex
Rewrite the experiment section
- let's be honest: {experiments->studies}
- rephrase the Tom and Jerry study to be more interesting

\subsubsection{Reasoning}

@@

\begin{table}
\begin{center}
\begin{tabular}{p{8cm}}
\hline
    {\bf Manipulation} \\
     {\tt attachobject}, {\tt basicgive}, {\tt basictake}, {\tt close\_gripper}, {\tt configure\_grippers}, {\tt grab\_gripper}, {\tt handover}, {\tt hide}, {\tt open\_gripper}, {\tt pick}, {\tt put}, {\tt put\_accessible}, {\tt release\_gripper}, {\tt show}, {\tt take} \\
\hline
    {\bf Gaze control} \\
     {\tt glance\_to}, {\tt look\_at}, {\tt sweep}, {\tt switch\_active\_stereo\_pair}, {\tt track}, {\tt cancel\_track} \\
\hline
    {\bf Navigation} \\
     {\tt carry}, {\tt follow}, {\tt cancel\_follow}, {\tt goto}, {\tt moveclose}, {\tt waypoints} \\
\hline
    {\bf Local navigation} \\
     {\tt dock}, {\tt rotate}, {\tt translate} \\
\hline
    {\bf Posture configuration} \\
     {\tt extractpose}, {\tt idle}, {\tt manipose}, {\tt movearm}, {\tt rest}, {\tt setpose}, {\tt settorso}, {\tt tuckedpose} \\
\hline
\end{tabular}
\end{center}
\caption{Main {\sc pyRobots} \emph{actions}, sorted by categories. These
actions are combined at run-time into higher-level \emph{tasks}.}

@@

\section{Experimental Support}
\label{sec:expe}

Our architecture has been deployed and tested in a number of studies on
several robotic platforms. Table~\ref{table|experiences} lists the most
significant ones, with their main focuses and reference publications for the
interested reader.

@@

\end{tabular}
\end{center}
\caption{Main experiments conducted with our cognitive architecture.}
\label{table|experiences}
\end{table*}

The two following sections present one such experiment (the {\it Interactive
Grounding I} scenario), and the synopsis of another one (the {\it Cleaning the
table} scenario).

The first one is focused on knowledge representation and verbal interaction:
the human asks for help to find and pack objects. In this experiment, robot
actions are limited to simple motions (like head tracking or predefined
pick-and-place).

The second one involves the {\sc Shary} execution controller and the HATP symbolic
task planner besides {\sc Oro} and {\sc Dialogs}. In this scenario, the human and the
robot try to cooperatively remove objects from a table. behaviours and motions
are fully planned.

\subsection{\emph{Moving to London} scenario}
\label{moving-london}

This first experiment is based on the following daily life situation: Tom and
Jerry are moving to London, so they are packing things in boxes. The scenario
takes places in the living-room, where Jido (our robot) is observing while they
move things here and there. To assess the reasoning abilities of the robot they
ask Jido for information. Ideally, the robot should
also perform actions when required (\eg hand an object when asking ``give
me...'').

Perception of objects is done through a tag-based system and humans are
detected through motion capture. The robot knowledge base is pre-loaded with
the \emph{{\sc Oro} Commonsense Ontology}.  We next describe two
situations where we can follow the internal robot's reasoning and the
interaction with the users.

\paragraph{Implicit disambiguation through visual perspective taking}

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.9\linewidth]{pt.jpg}
\caption{Asking the robot to hand over a tape on the table.}
  \label{fig|vpt}
\end{figure}

Tom enters the room while carrying a big box (Figure~\ref{fig|vpt}). He
approaches the table and asks Jido to hand him the video tape: ``Jido, can
you give me the video tape''. The \textsc{Dialogs} module queries the ontology to
identify the object the human is referring to: \stmt{?obj type VideoTape}.

@@

However, only one is visible for Tom (the one on the
table). Thus, although there is an ambiguity from the robot's perspective
(since it can see both video tapes), based on the perspective of its human
partner it infers that Tom is referring to the video tape on the table, and not
the one inside the box which is not visible for him.

Since only one object remains (\ie the referent that \concept{VideoTape} stands
for is not ambiguous anymore), the robot infers that the human refers to it and
eventually execute the command, \ie give it to the human. Alternatively, the
robot first asks the human for confirmation (``was \concept{videoTape1} the
object you had in mind?'') before proceeding with the action.
Table~\ref{table|ptbeliefs} lists the robot's beliefs about itself and its human
partner involved in this situation.

\begin{table}[ht!]
\begin{center}
    \subfigure{
\begin{tabular}{l}
\hline
Robot's beliefs about itself (\emph{robot's model}):\\
\hline
  \hspace{0.7cm}\stmt{videoTape1 type VideoTape}\\
  \hspace{0.7cm}\stmt{videoTape1 isOn table}\\
  \hspace{0.7cm}\stmt{videoTape1 isVisible true}\\
  \hspace{0.7cm}\stmt{videoTape2 type VideoTape}\\
  \hspace{0.7cm}\stmt{videoTape2 isIn cardBoardBox}\\
  \hspace{0.7cm}\stmt{videoTape2 isVisible true}\\
\hline
\end{tabular}
}\hspace{1em}%
    \subfigure{
\begin{tabular}{l}
\hline
Robot's beliefs about Tom (\emph{Tom's model}):\\
\hline
  \hspace{0.7cm}\stmt{videoTape1 type VideoTape}\\
  \hspace{0.7cm}\stmt{videoTape1 isOn table}\\
  \hspace{0.7cm}\stmt{videoTape1 isVisible true}\\
  \hspace{0.7cm}\stmt{videoTape2 type VideoTape}\\
  \hspace{0.7cm}\stmt{videoTape2 isIn cardBoardBox}\\
  \hspace{0.7cm}\stmt{videoTape2 isVisible false}\\
 \hline
\end{tabular}
}
\end{center}
\caption{Robot's beliefs about itself and its human partner.}
\label{table|ptbeliefs}
\end{table}

\paragraph{Explicit disambiguation through verbal interaction and gestures}

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.9\linewidth]{inTheBox2.jpg}
\caption{Jerry asks Jido for the content of the box by pointing at it.}
  \label{fig|pointing}
\end{figure}

In this situation, Jerry enters the living room without knowing where Tom had
placed the video tapes (Figure~\ref{fig|pointing}). So he first asks Jido:
``What's in the box?''. Before the robot can answer the question it has to
figure out which box Jerry is talking about. Similar to the previous situation,
there are two available boxes:

@@

However both are visible and the cognitive ambiguity resolution cannot be
applied. The only option is to ask Jerry which box he is referring to: ``Which
box, the toolbox or the cardboard box?'' Jerry could answer the
question. Instead, he decides to point at it while indicating: ``This box''
(Figure~\ref{fig|pointing}). The robot's perception identifies the {\tt
cardBoardBox} as being pointed at and looked at by the human and updates the
ontology with this new information using a rule available in the commonsense
ontology ({\tt \textbf{pointsAt}(?ag, ?obj) $\land$ \textbf{looksAt}(?ag, ?obj) $\to$
\textbf{focusesOn}(?ag, ?obj)}). The \textsc{Dialogs} module is then able to merge both
sources of information, verbal (``this'') and deictic to distinguish the box
Jerry refers to:
\begin{center}
\begin{tabular}{l}
\stmt{Jerry pointsAt carboardBox}\\
\stmt{Jerry looksAt carboardBox}\\
$\to$ \stmt{Jerry focusesAt carboardBox}\\
\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = [cardBoardBox]}
\end{tabular}
\end{center}

Finally, the \textsc{Dialogs} queries the ontology about the content of the box
and the question can be answered: ``Jido-E''. Note that the object's label is
used instead of its ID. This way we enhance interaction using familiar names
given by the users:

\begin{center}
\stmt{?obj isIn cardBoardBox}\\
\hspace{0.7cm}$\Rightarrow$ \concept{?obj = videoTape2}\\
\end{center}

At this point Jerry wants to know where the other tape is, and that is exactly
what he asks Jido: ``And where is the other tape?''. In this situation, the
\textsc{Dialogs} module is able to interpret that Jerry is not referring to the
video which they were just talking about, but to the other one:

@@

Since there is only one possible ``other'' video (there are only two videos in
the scene), it can directly answer Jerry: ``The other tape is on the table and
next to the toolbox.''

>>>>

\subsubsection{Reasoning}
\label{reasoning}

@@

\begin{table}[ht!]
\begin{center}
\begin{tabular}{p{0.8\columnwidth}}
\hline
    {\bf Manipulation} \\
     {\tt attachobject}, {\tt basicgive}, {\tt basictake}, {\tt close\_gripper}, {\tt configure\_grippers}, {\tt grab\_gripper}, {\tt handover}, {\tt hide}, {\tt open\_gripper}, {\tt pick}, {\tt put}, {\tt put\_accessible}, {\tt release\_gripper}, {\tt show}, {\tt take} \\
\hline
    {\bf Gaze control} \\
     {\tt glance\_to}, {\tt look\_at}, {\tt sweep}, {\tt switch\_active\_stereo\_pair}, {\tt track}, {\tt cancel\_track} \\
\hline
    {\bf Navigation} \\
     {\tt carry}, {\tt follow}, {\tt cancel\_follow}, {\tt goto}, {\tt moveclose}, {\tt waypoints} \\
\hline
    {\bf Local navigation} \\
     {\tt dock}, {\tt rotate}, {\tt translate} \\
\hline
    {\bf Posture configuration} \\
     {\tt extractpose}, {\tt idle}, {\tt manipose}, {\tt movearm}, {\tt rest}, {\tt setpose}, {\tt settorso}, {\tt tuckedpose} \\
\hline
\end{tabular}
\end{center}
\caption{Main {\sc pyRobots} \emph{actions}, sorted by categories. These
actions are combined at run-time into higher-level \emph{tasks}.}

@@

\section{Support Studies}
\label{sec:expe}

Our architecture has been deployed and tested in a number of studies on
several robotic platforms. Table~\ref{table|experiences} lists the most
significant ones, with their main focuses and reference publications.

@@

\end{tabular}
\end{center}
\caption{Main studies conducted with our cognitive architecture.}
\label{table|experiments}
\end{table*}

We briefly introduce two of them in order to illustrate in a practical way
different aspects of the architecture that were previously presented.  The first
one is focused on knowledge representation and verbal interaction: the human
asks for help to find and pack objects ({\it Interactive Grounding I} in
Table~\ref{table|experiments}). The second one ({\it Cleaning the table})
involves the {\sc Shary} execution controller and the HATP symbolic task planner
besides {\sc Oro} and {\sc Dialogs}. In this scenario, the human and the robot
try to cooperatively remove objects from a table, and behaviours and motions are
fully planned and executed.

\subsection{Interactive Grounding}
\label{moving-london}

This first study is based on the following scenario: Tom and Jerry are moving to
London, so they are packing things in boxes. \emph{Jido}, a dual-arm
manipulator, is observing while they move things here and there
(Figure~\ref{fig|vpt}).  They ask questions to Jido to know where certain
objects are located (\emph{Where is the white tape? What is behind the toolbox?}
etc.). This study focuses on multi-modal ,interactive grounding only, and
the robot does not perform any action.

Objects are perceived solely through 2D fiducial markers sticked on them, and humans are
tracked through motion capture. The robot knowledge base is initialised with
the {\sc Oro} commonsense ontology.  We next describe two
situations where we can follow the internal robot's reasoning and the
interaction with the users.

\paragraph{Implicit Disambiguation Through Visual Perspective Taking}

\begin{figure}[!ht]
  \centering
  \subfigure [Interactive grounding in a cluttered environment.]{
  \label{fig|vpt}
  \includegraphics[width=0.48\linewidth]{pt.jpg}
}%
\subfigure [Disambiguation through pointing.]{
  \label{fig|pointing}
  \includegraphics[width=0.48\linewidth]{inTheBox2.jpg}
}
\end{figure}

Tom enters the room while carrying a big box (Figure~\ref{fig|vpt}). He
approaches the table and asks Jido to hand him the video tape: ``Jido, can
you give me the video tape''. The \textsc{Dialogs} module processes this
sentence, and queries the ontology to
identify the object the human is referring to: \stmt{?obj type VideoTape}.

@@

However, only one is visible for Tom (the one on the table). Although there is
an ambiguity from the robot's perspective, the human referred to the tape using
the definite article \emph{the}: this is interpreted by the natural language
processor as the human referring to a precise object, in that case, the visible
one.\footnote{Other heuristics are available to the {\sc Dialogs} module: for
instance, if a tape had been recently mentioned in the dialogue, this instance
would have been selected instead as the referent.}

%Since only one object remains (\ie the referent that \concept{VideoTape} stands
%for is not ambiguous any more), the robot infers that the human refers to it and
%eventually execute the command, \ie give it to the human. Alternatively, the
%robot first asks the human for confirmation (``was \concept{videoTape1} the
%object you had in mind?'') before proceeding with the action.
%Table~\ref{table|ptbeliefs} lists the robot's beliefs about itself and its human
%partner involved in this situation.
%
%\begin{table}[ht!]
%\begin{center}
%    \subfigure{
%\begin{tabular}{l}
%\hline
%Robot's beliefs about itself (\emph{robot's model}):\\
%\hline
%  \hspace{0.7cm}\stmt{videoTape1 type VideoTape}\\
%  \hspace{0.7cm}\stmt{videoTape1 isOn table}\\
%  \hspace{0.7cm}\stmt{videoTape1 isVisible true}\\
%  \hspace{0.7cm}\stmt{videoTape2 type VideoTape}\\
%  \hspace{0.7cm}\stmt{videoTape2 isIn cardBoardBox}\\
%  \hspace{0.7cm}\stmt{videoTape2 isVisible true}\\
%\hline
%\end{tabular}
%}\hspace{1em}%
%    \subfigure{
%\begin{tabular}{l}
%\hline
%Robot's beliefs about Tom (\emph{Tom's model}):\\
%\hline
%  \hspace{0.7cm}\stmt{videoTape1 type VideoTape}\\
%  \hspace{0.7cm}\stmt{videoTape1 isOn table}\\
%  \hspace{0.7cm}\stmt{videoTape1 isVisible true}\\
%  \hspace{0.7cm}\stmt{videoTape2 type VideoTape}\\
%  \hspace{0.7cm}\stmt{videoTape2 isIn cardBoardBox}\\
%  \hspace{0.7cm}\stmt{videoTape2 isVisible false}\\
% \hline
%\end{tabular}
%}
%\end{center}
%\caption{Robot's beliefs about itself and its human partner.}
%\label{table|ptbeliefs}
%\end{table}
%
\paragraph{Explicit Disambiguation Through Verbal Interaction and Gestures}

In this second situation, Jerry enters the living room without knowing where Tom had
placed the video tapes (Figure~\ref{fig|pointing}). So he first asks Jido:
``What's in the box?''. Before the robot can answer the question it has to
figure out which box Jerry is talking about. Similar to the previous situation,
two boxes are visible:

@@

However both are also visible to the human and the previous ambiguity resolution
procedure can not be applied. The robot generates a question (using the
\emph{Discrimination} algorithm that we introduced at section~\ref{reasoning})
and asks Jerry which box he is referring to: ``Which box, the toolbox or the
cardboard box?'' Jerry can answer the question, but he instead decides to point
at it: ``This box'' (Figure~\ref{fig|pointing}). {\sc Spark} identifies the {\tt
cardBoardBox} as being pointed at and looked at by the human and updates the
ontology with this new information using a rule available in the common-sense
ontology \stmt{pointsAt(?ag, ?obj) $\land$ looksAt(?ag, ?obj) $\to$
focusesOn(?ag, ?obj)}. The \textsc{Dialogs} module is then able to merge both
sources of information, verbal (``this'') and deictic to distinguish the box
Jerry refers to:

\begin{center} 
    \begin{tabular}{l} 
        \stmt{Jerry pointsAt carboardBox}\\ 
        \stmt{Jerry looksAt carboardBox}\\ 
        $\to$ \stmt{Jerry focusesAt carboardBox}\\ 
        \hspace{0.7cm}$\Rightarrow$ {\tt ?obj = [cardBoardBox]}
    \end{tabular} 
\end{center}

Finally, the \textsc{Dialogs} queries the ontology about the content of the box
and the question can be answered: ``Wall-E'' (the label of the object, stored in the
ontology, is used to interact with the user).

%\begin{center}
%\stmt{?obj isIn cardBoardBox}\\
%\hspace{0.7cm}$\Rightarrow$ \concept{?obj = videoTape2}\\
%\end{center}

At this point Jerry wants to know where is the other tape: ``And where is the
other tape?''. \textsc{Dialogs} processes this sentence using the
\concept{differentFrom} OWL predicate:

@@

Since there is only one possible ``other'' video, no specific disambiguation is
required, and the robot verbalises the result of the query to the knowledge
base: ``The other tape is on the table and next to the toolbox.''
